---
title: "Homework 5"
output: github_document
date: "2023-11-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(rvest)
library(readr)
set.seed(1)
```

## PROBLEM 1

The code chunk below reads in the Washington Post homicide data, and creates a new variable `city_state` that combines the city and state information.
```{r}
urlfile = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicide = read_csv(url(urlfile)) |>
          janitor::clean_names() |>
          mutate(
            city_state = paste(city, state, sep = ", ")
          )
```

There are 52179 observations and 13 variables in the homicide dataset detailing homicides in 50 states over a 10 year period (January 2007 to December 2017). Variables include date of homicide (`reported_date`), name of victim (`victim_last` and `victim_first`), victim demographics (`victim_race`, `victim_sex`, and `victim_age`), where the crime occurred (`city`, `state`, `city_state`, `lat`, `long`), as well as the status of the case - whether the case is open or closed and if there has been an arrest (`disposition`).

